Towards a Verified Bzip2

### 2019-10-06 draft_paper/draft_paper.tex
Fix link
Compile succesfully(ish) with acmart

\documentclass[11pt]{thesis}

\title{Towards a Verified Bzip2: A Functional Specification for Burrows--Wheeler}
\author{Jake Waksbaum}
\date{May 6, 2019}

@@

\pagenumbering{gobble}

\begin{titlepage}
  \centering \Large{\textbf{Towards a Verified Bzip2: \\ A Functional Specification for Burrows--Wheeler
  }} \vspace{10mm}

  \Large\textit{Jake Waksbaum} \\
  \vspace{25mm} \large{Junior Two-semester Independent Work \\ Department of
    Computer Science \\ Princeton University \\ Fall 2018--Spring 2019\\}
  \vspace{8mm}
  \begin{figure}[!h]
    \centering
    \includegraphics[scale=0.1]{seal.png}
  \end{figure}
  \vspace{1mm}
  \normalsize{Advised by Andrew W. Appel (COS)}\\
  \normalsize{Second Reader: David Walker (COS)}

  \vfill
  \pledge{Jake Waksbaum}
\end{titlepage}

\clearpage

\section*{Acknowledgements}

Thanks to Professor Andrew Appel for being extremely generous with his
time and his guidance. He was always available for questions, and
allowed me to focus my work on the areas I found interesting.

I'd like to thank Benjamin Huang, Yael Marans, and Eli Waksbaum for
their help and support.

\clearpage

\tableofcontents
\clearpage

\pagenumbering{arabic}

@@

Every definition and theorem in this paper corresponds directly to a
definition or theorem in our Coq sources, which can be found at
\url{https://github.com/jbaum98/verifed_bzip2/tree/2019-05-06}.
Definitions are copied verbatim. Theorems are rephrased for clarity,
with the name of the corresponding Coq theorem given in parentheses.
The only exceptions are \var{sec:opt} and places where we give
hypothetical definitions to provide an explanation.

@@

\begin{theorem}[iter\_nth]
  For all $f$, $n$, $z$, and $i < n$,
  \begin{equation*}
    (\var{iter}~f~n~z)_i = f^i~z
  \end{equation*}
\end{theorem}
\begin{proof}
  The theorem follows by induction on $i$. The inductive case is
  \begin{align*}
       (\var{iter}~f~(n+1)~z)_{i+1} \\
    &= (z :: \var{iter}~f~n~(f~z))_{i+1} \\
    &= (\var{iter}~f~n~(f~z))_i \\
    &= f^i~(f~z) \\
    &= f^{i+1}~z \qedhere
  \end{align*}
\end{proof}

@@

This implementation, including the optimizations mentioned in
\cref{sec:opt}, can be used as a functional specification to for a
verified implementation of bzip2. Of course, it would also be
necessary to have functional specifications for the other phases of
bzip2, such as Run-Length Encoding, Move-to-Front Encoding, and
arithmetic coding. As luck would have it, Bird has already analyzed
arithmetic coding \cite{pearls}, and the other two phases are
relatively simple.
\clearpage

@@

\clearpage

\nocite{*}
\printbibliography{}

>>>>

\documentclass[sigplan,10pt,anonymous,review]{thesis}

\acmConference[PL'18]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2018}{New York, NY, USA}
\acmYear{2020}
\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%\citestyle{acmauthoryear}  %% For author/year citations
%\citestyle{acmnumeric}     %% For numeric citations
%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
                            %% sorting of references within a single citation;
                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
                            %% rendered as [14,5,2] rather than [2,5,14].
%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
                            %% compression of sequential references within a
                            %% single citation;
                            %% e.g., \cite{Baker12,Baker14,Baker16}
                            %% rendered as [2,3,4] rather than [2-4].

@@

\title{Towards a Verified Bzip2}
\subtitle{A Functional Specification for Burrows--Wheeler}

\author{Jake Waksbaum}
\affiliation{
  \department{Department of Computer Science}
  \institution{Princeton University}
  \city{Princeton}
  \state{New Jersey}
  \postcode{08540}
  \country{USA}
}
\email{waksbaum@princeton.edu}

\author{Andrew Appel}
\affiliation{
  \department{Department of Computer Science}
  \institution{Princeton University}
  \city{Princeton}
  \state{New Jersey}
  \postcode{08540}
  \country{USA}
}
\email{appel@princeton.edu}

\begin{abstract}
Text of abstract \ldots.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML} <ccs2012>
<concept>
<concept_id>10003752.10003809.10010031.10002975</concept_id>
<concept_desc>Theory of computation~Data compression</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003752.10003809.10010031.10010033</concept_id>
<concept_desc>Theory of computation~Sorting and searching</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10010940.10010992.10010998</concept_id>
<concept_desc>Software and its engineering~Formal methods</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Data compression}
\ccsdesc[500]{Theory of computation~Sorting and searching}
\ccsdesc[500]{Software and its engineering~Formal methods}
%% End of generated code

\keywords{Formal methods, Sorting and seraching, Data compression}

\maketitle

@@

Every definition and theorem in this paper corresponds directly to a
definition or theorem in our Coq sources, which can be found at
\url{https://github.com/jbaum98/verified_bzip2/tree/2019-05-06}.
Definitions are copied verbatim. Theorems are rephrased for clarity,
with the name of the corresponding Coq theorem given in parentheses.
The only exceptions are \var{sec:opt} and places where we give
hypothetical definitions to provide an explanation.

@@

\begin{theorem}[iter\_nth]
  For all $f$, $n$, $z$, and $i < n$,
  \begin{equation*}
    (\var{iter}~f~n~z)_i = f^i~z
  \end{equation*}
\end{theorem}
\begin{proof}
  The theorem follows by induction on $i$. The inductive case is
  \begin{align*}
       (\var{iter}~f~(n+1)~z)_{i+1}
    &= (z :: \var{iter}~f~n~(f~z))_{i+1} \\
    &= (\var{iter}~f~n~(f~z))_i \\
    &= f^i~(f~z) \\
    &= f^{i+1}~z \qedhere
  \end{align*}
\end{proof}

@@

This implementation, including the optimizations mentioned in
\cref{sec:opt}, can be used as a functional specification to for a
verified implementation of bzip2. Of course, it would also be
necessary to have functional specifications for the other phases of
bzip2, such as Run-Length Encoding, Move-to-Front Encoding, and
arithmetic coding. As luck would have it, Bird has already analyzed
arithmetic coding \cite{pearls}, and the other two phases are
relatively simple.

\begin{acks}
Thanks to Professor Andrew Appel for being extremely generous with his
time and his guidance. He was always available for questions, and
allowed me to focus my work on the areas I found interesting.

I'd like to thank Benjamin Huang, Yael Marans, and Eli Waksbaum for
their help and support.
\end{acks}

\bibliography{draft_paper}

@@
